Deep Learning

### 2015-01-11 deepreview.tex
cover reference and a little explanation

This review is organized chronlogically starting with the introduction of
the backpropagation algorithm in the 1980's. After describing some of its
early successes, we explain the actual and perceived limitations that
prevented it from living up to its early promise for so many years and how
these limitations were eventually overcome.

@@

However, even with good hand-crafted features, many interesting tasks
cannot be solved by linear classifiers. A common method to get around this
problem is to transform the feature vector into a higher dimensional vector
through a non-linear mapping, exploiting the fact that collections of
vectors are more likely to be easily spearated when they are spread around
a high-dimensional space~\cite{cover}. One clever way to do so is to use a
``kernel function'' that measures the similarity between two vectors.
Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input
vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
optimize the choice of training inputs used to define features and the
class-specific weights associated with the features~\cite{Boser92}. The training
inputs used to define features are called ``support vectors'' and Support
Vector Machines~\citep{Boser92} have been very successful, especially for
tasks with a relatively small number of training examples in which features
are individually informative. They do not, however, perform well for tasks
such as object recognition in which the input space is high-dimensional and
the input-output mapping is so complex that computing similarities with a
simple kernel function is ineffective.

@@

Prior to the introduction of neural language
models~\citet{BenDucVin01-short} the standard approach to statistical
modeling of language was based on counting frequencies of occurences of
short symbol sequences (called n-grams). One disadvantage is that it
taking into account context beyond a handful of words would require
training corpora of exponentially growing size (with respect to the
desired context size), since the number of configurations of $n$
words from a vocabulary of size $V$ grows as $V^n$. Related to this
problem and the very notion of symbol introduced above, n-grams
do not generalize across the many semantically related sequences 
of words (for example the many ways of expressing roughly the same thing).
Instead, neural language models learn to associate each word to a
distributed representation (a vector) such that semantically related
words end up close to each other in that space, as illustrated in
Figure~\ref{fig:word-embeddings}. To obtain that effect, one
{\em jointly learns} a representation for each word as well as
a function that computes a quantity of interest that is relevant
to a language related task, such as predicting the next word
in a sequence (neural language models) or predicting the sequence
of translated words given the sequence of source words, i.e., neural
machine translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

@@

When translating between languages, the hidden units of the decoder network
that generates possible translations are initalized with a ``thought
vector'' which is the final hidden state of an encoder network.  But it is
also possible to initialize a decoder network using a linear transformation
of the last hidden layer of a deep convolutonal network that has already
been trained to recognize objects in images. The decoder network can then
be trained by BPTT on a dataset of captioned
images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
captions (see Figure~\ref{fig:caption-generation}).

>>>>

This review is organized chronologically starting with the introduction of
the backpropagation algorithm in the 1980's. After describing some of its
early successes, we explain the actual and perceived limitations that
prevented it from living up to its early promise for so many years and how
these limitations were eventually overcome.

@@

However, even with good hand-crafted features, many interesting tasks
cannot be solved by linear classifiers. A common method to get around this
problem is to transform the feature vector into a higher dimensional vector
through a non-linear mapping, exploiting the fact that collections of
vectors are more likely to be easily separated when they are spread around
a high-dimensional space~\cite{Cover65}. In the limit when there are at least
as many dimensions as examples, a linear classifier can always separate the
training examples, and one clever way to obtain a very high-dimensional space is to use a
``kernel function'' that measures the similarity between two vectors.
Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input
vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
optimize the choice of training inputs used to define features and the
class-specific weights associated with the features~\cite{Boser92}. The training
inputs used to define features are called ``support vectors'' and Support
Vector Machines~\citep{Boser92} have been very successful, especially for
tasks with a relatively small number of training examples in which features
are individually informative. They do not, however, perform well for tasks
such as object recognition in which the input space is high-dimensional and
the input-output mapping is so complex that computing similarities with a
simple kernel function is ineffective.

@@

Prior to the introduction of neural language
models~\citet{BenDucVin01-short} the standard approach to statistical
modeling of language was based on counting frequencies of occurrences of
short symbol sequences (called n-grams). One disadvantage is that it
taking into account context beyond a handful of words would require
training corpora of exponentially growing size (with respect to the
desired context size), since the number of configurations of $n$
words from a vocabulary of size $V$ grows as $V^n$. Related to this
problem and the very notion of symbol introduced above, n-grams
do not generalize across the many semantically related sequences 
of words (for example the many ways of expressing roughly the same thing).
Instead, neural language models learn to associate each word to a
distributed representation (a vector) such that semantically related
words end up close to each other in that space, as illustrated in
Figure~\ref{fig:word-embeddings}. To obtain that effect, one
{\em jointly learns} a representation for each word as well as
a function that computes a quantity of interest that is relevant
to a language related task, such as predicting the next word
in a sequence (neural language models) or predicting the sequence
of translated words given the sequence of source words, i.e., neural
machine translation~\citep{Bahdanau-et-al-arxiv2014,Sutskever-et-al-NIPS2014}.

@@

When translating between languages, the hidden units of the decoder network
that generates possible translations are initialized with a ``thought
vector'' which is the final hidden state of an encoder network.  But it is
also possible to initialize a decoder network using a linear transformation
of the last hidden layer of a deep convolutonal network that has already
been trained to recognize objects in images. The decoder network can then
be trained by BPTT on a dataset of captioned
images~\cite{NIC,StanfordNIC,more?} and it produces remarkably good
captions (see Figure~\ref{fig:caption-generation}).
